Continuous developmental change can explain discontinuities in word learning

### 2019-02-01 cogsci/cogsci.Rmd
final version cogsci

Subsequent findings have suggested otherwise. On the one hand, 14-month-olds---who typically fail in the original tas---succeed when an easier testing method is used, even under the same learning conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

@@

$$ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) $$
$$ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) $$
Finally, we assume there to be one-to-one mappings between concepts and labels, and that observers have successfully learned these mappings during the exposure phase:
$$
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
$$

@@

We used Equation 1 to obtain the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\\% confidence intervals."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

Experimental results are shown in Figure 6 (top). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings. For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that performance improves with age. In addition, we found two-way interactions between sound distance and age ($\beta =$ `r s_age_coef`, $p$ `r p_val(s_age_p)`) and between object distance and age ($\beta =$ `r c_age_coef`, $p$ `r p_val(c_age_p)`).

@@

We next fit our model (using Equation 2) to the participants' responses in each age group.  The values of $\Delta s$ and $\Delta o$ were set based on data from the similarity judgment task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom for each group, i.e., $\sigma_C$ and $\sigma_L$.  Figure \ref{fig:data_all} (bottom) shows the predictions. The model captures the qualitative patterns in both age groups: starting from a low accuracy recognition when both the sound and object distances are small, the model correctly predicts an increase in accuracy when either the sound distance or the object distance increases. Further, accuracy is correctly predicted to be maximal when both the sound and object distances are high.

@@

The model made a new prediction to which which we tested experimentally: Learning similar words is not only modulated by the similarity of their phonological forms, but also by the visual similarity of their semantic referents. More generally, since visual similarity is an early organizing feature in the semantic domain [e.g.,@wojcik2013], our finding suggests that children may prioritize the acquisition of words that are quite distant in the semantic space.  This suggestion is supported by recent findings based on the investigation of early vocabulary growth [@engelthaler2017; @sizemore2018].

@@

In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. Our account suggests that the developmental data can be explained based on a continuous process operating over similar representations across development, suggesting developmental continuity. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models such as the one proposed here can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes, and the extent to which they reflect the granularity of the researchers' own measurement tools.

>>>>

Subsequent findings have suggested otherwise. On the one hand, 14-month-olds---who typically fail in the original task---succeed when an easier testing method is used, even under the same learning conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

@@

$$ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) $$
$$ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) $$
Finally, we assume there to be one-to-one mappings between concepts and labels and that observers have successfully learned these mappings during the exposure phase:
$$
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
$$

@@

We used Equation 1 and the probability distributions defined above to obtain the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\\% confidence intervals."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

Experimental results are shown in Figure 6 (solid lines). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings. For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that performance improves with age. In addition, we found two-way interactions between sound distance and age ($\beta =$ `r s_age_coef`, $p$ `r p_val(s_age_p)`) and between object distance and age ($\beta =$ `r c_age_coef`, $p$ `r p_val(c_age_p)`).

@@

We next fit our model (using Equation 2) to the participants' responses in each age group.  The values of $\Delta s$ and $\Delta o$ were set based on data from the similarity judgment task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom for each group, i.e., $\sigma_C$ and $\sigma_L$.  Figure \ref{fig:data_all} (dashed lines) shows the predictions. The model captures the qualitative patterns in both age groups: starting from a low accuracy recognition when both the sound and object distances are small, the model correctly predicts an increase in accuracy when either the sound distance or the object distance increases. Further, accuracy is correctly predicted to be maximal when both the sound and object distances are high.

@@

The model made a new prediction to which we tested experimentally: Learning similar words is not only modulated by the similarity of their phonological forms, but also by the visual similarity of their semantic referents. More generally, since visual similarity is an early organizing feature in the semantic domain [e.g.,@wojcik2013], our finding suggests that children may prioritize the acquisition of words that are quite distant in the semantic space.  This suggestion is supported by recent findings based on the investigation of early vocabulary growth [@engelthaler2017; @sizemore2018].

@@

In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. Our account suggests that the developmental data can be explained based on a continuous process operating over similar representations across development, suggesting developmental continuity. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models, such as the one proposed here, can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes and the extent to which they reflect the granularity of the researchers' own measurement tools.

### 2019-02-01 cogsci/cogsci.tex
final version cogsci

Subsequent findings have suggested otherwise. On the one hand,
14-month-olds---who typically fail in the original tas---succeed when an
easier testing method is used, even under the same learning conditions
(Yoshida, Fennell, Swingley, \& Werker, 2009). They also succeed when
uncertainty is mitigated via disambiguating cues (e.g., Thiessen, 2007).
On the other hand, adults show patterns of learning similar to those
shown by 14-month-olds when the task is more challenging and when
similarity between words increases (Pajak, Creel, \& Levy, 2016; White,
Yee, Blumstein, \& Morgan, 2013).

@@

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]
\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \] Finally, we assume
there to be one-to-one mappings between concepts and labels, and that
observers have successfully learned these mappings during the exposure
phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\]

@@

We used Equation 1 to obtain the exact analytical expression for the
probability of accurate responses \(p(o_T | s)\) (target object \(o_T\)
given a sound \(s\)) in the simple case of two-alternative forced choice
in the testing phase of our experimental task:

@@

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\% confidence intervals.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

The model made a new prediction to which which we tested experimentally:
Learning similar words is not only modulated by the similarity of their
phonological forms, but also by the visual similarity of their semantic
referents. More generally, since visual similarity is an early
organizing feature in the semantic domain (e.g., Wojcik \& Saffran,
2013), our finding suggests that children may prioritize the acquisition
of words that are quite distant in the semantic space. This suggestion
is supported by recent findings based on the investigation of early
vocabulary growth (Engelthaler \& Hills, 2017; Sizemore, Karuza, Giusti,
\& Bassett, 2018).

>>>>

Subsequent findings have suggested otherwise. On the one hand,
14-month-olds---who typically fail in the original task---succeed when
an easier testing method is used, even under the same learning
conditions (Yoshida, Fennell, Swingley, \& Werker, 2009). They also
succeed when uncertainty is mitigated via disambiguating cues (e.g.,
Thiessen, 2007). On the other hand, adults show patterns of learning
similar to those shown by 14-month-olds when the task is more
challenging and when similarity between words increases (Pajak, Creel,
\& Levy, 2016; White, Yee, Blumstein, \& Morgan, 2013).

@@

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]
\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \] Finally, we assume
there to be one-to-one mappings between concepts and labels and that
observers have successfully learned these mappings during the exposure
phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\]

@@

We used Equation 1 and the probability distributions defined above to
obtain the exact analytical expression for the probability of accurate
responses \(p(o_T | s)\) (target object \(o_T\) given a sound \(s\)) in
the simple case of two-alternative forced choice in the testing phase of
our experimental task:

@@

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\% confidence intervals.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

The model made a new prediction to which we tested experimentally:
Learning similar words is not only modulated by the similarity of their
phonological forms, but also by the visual similarity of their semantic
referents. More generally, since visual similarity is an early
organizing feature in the semantic domain (e.g., Wojcik \& Saffran,
2013), our finding suggests that children may prioritize the acquisition
of words that are quite distant in the semantic space. This suggestion
is supported by recent findings based on the investigation of early
vocabulary growth (Engelthaler \& Hills, 2017; Sizemore, Karuza, Giusti,
\& Bassett, 2018).
